Prestige bias in cultural evolutionary dynamics

### 2023-11-07 ms/ms.tex
revised introduction and discussion following rev1; fixed bibitems

% Title page
\title{Influence bias in cultural evolutionary dynamics}
% Authors
\renewcommand\Affilfont{\small}

\author[1,3]{Saar Egozi}
\author[2,3,$\dagger$]{Yoav Ram}

\affil[1]{School of Computer Science, Reichman University, Herzliya, Israel}
\affil[3]{School of Zoology, Faculty of Life Sciences, Tel Aviv University, Tel Aviv, Israel}
\affil[3]{Sagol School of Neuroscience, Tel Aviv University, Tel Aviv, Israel}
\affil[$\dagger$]{Corresponding author: yoav@yoavram.com}

\begin{document}
\maketitle

\begin{abstract}
{If the traits of more successful individuals are more likely to be adopted, the resulting cultural transmission is described as being success-biased. Prestige may be used as a proxy for success.}
Here, we model influence bias as a combination of indirect success bias and prestige bias, the latter meaning that the choice of a role-model depends on the number of individuals that have already copied  that role-model.
{The} effect of prestige on cultural evolutionary dynamics {is analyzed} using mathematical analysis and stochastic simulations.
{Analytic} approximations to the stochastic role-model choice process {facilitate the} mathematical analysis and reduce the computational complexity of simulations. {Approximations are given}  to the fixation probability and the fixation time of an invading cultural trait in different environments. 
We show that indirect success bias effectively plays the role of natural selection, whereas prestige bias effectively plays the role of genetic drift.	
Influence bias{, which may be strong in communities where social media are prevalent,}  also accelerates the evolutionary dynamics, as can be expected in a rich-{get}-richer process.
\end{abstract}

@@

\section*{Introduction}

Cultural transmission {of attitudes, preferences, beliefs, norms, and behaviors} may combine vertical transmission, in which parents transmit to their offspring; oblique transmission, in which adults (teachers, leaders, and even strangers) transmit to unrelated offspring; and horizontal transmission, in which {individuals} from the same age cohort transmit to one another \citep{transmissionVectorsBook}.
It has been demonstrated that non-vertical cultural transmission can maintain maladaptive traits, which can be beneficial in changing environments \citep{changeEnv,Lehmann2009}.

Transmission biases may cause a cultural trait to have a {higher rate of transmission than} its frequency in the population. \textit{Success bias} occurs when individuals prefer to copy from role models that demonstrate success in some activity, such as fishing, growing yams, using medicinal plants \citep{fijian_social_bias}, or hunting \citep{Mesoudi2008}, and it can increase the probability of learning a {trait that is present in those successful individuals} \citep{Borofsky2022}. 
Indeed, in a tournament between learning strategies  \citep{strategiesPaper}, most {winning} strategies included a mixture of success-biased social learning and individual learning, implying that success-biased learning is a good strategy, but that by itself it is not enough to best other strategies, even when success is measured accurately.

{Boyd and Richerson} \citep[Ch.~5]{evolutionBook} suggested that the {assessment} of success can be divided to three categories: \emph{direct bias}, \emph{indirect bias} and \emph{frequency-dependent bias}.
{Direct} bias occurs when one phenotype is more attractive than other phenotypes, and is evaluated by \emph{directly} testing the trait.
For example, an individual observing a ping-pong match can {try} the observed paddle grips to determine which grip is better.
Frequency-dependent bias occurs when the probability {of copying} a phenotype is higher or lower than the frequency of the phenotype among demonstrators. 
For example, suppose the common paddle grip is used by 60\% of the demonstrators; if the this grip is adopted by 80\% of copiers, then transmission is under positive frequency bias, also called \emph{conformity}; if it is adopted by 40\% of copiers, then transmission is under negative frequency bias, or \emph{{anti}-conformity} \citep{conformism}. 
The effects of conformity and {anti}-conformity on cultural evolution have been studied with both models \citep{anticonformity,Denton2021,conformity} and experiments \citep{negativeFrequency}.
Indirect bias occurs when a copier uses some observed phenotype to evaluate the attractiveness of a potential role-model. 
For example, an observer may copy the paddle grip of the ping-pong player who scored more points in the match, thus indirectly evaluating the grip by the points scored.
However, this may cause mismatches between the copied trait and the rest of the cultural or genetic repertoire of the individual \citep{Kolodny2022}.
Furthermore, {Boyd and Richerson} \citep[Ch. 8]{evolutionBook} {suggest} that maladaptive traits may spread widely in a population if indirect {bias is} strong enough, e.g., by a runaway process caused by a cultural equivalent of sexual selection \citep{sexualSelectionBook}.
Indeed, helping behaviors can evolve due to horizontal transmission bias even without any benefit to the recipient, or when the benefit is much larger than the cost \citep{cooperation}.

Henrich and Broesch \citep{fijian_social_bias}  studied such indirect success biases, which they call cross-domain success bias or \emph{prestige bias} (e.g., great fishermen may be chosen as role-models for growing yams). They {suggested} that such biases, over generations, can lead to cultural adaptations, and that although prestige can lead to maladaptive traits spreading in the population, it can also accelerate the spread of adaptive traits
\citep[Ch.~8]{evolutionBook}. {Prestige bias may be} more common in humans than success {bias} \citep{complexityPaper}, and prestige is often mentioned in the cultural evolution literature, {although} there are few models of {it}.

A broader definition of prestige is the ``widespread respect and admiration felt for someone or something on the basis of a perception of their achievements or quality'' (New Oxford American Dictionary).
Indeed, Chudek et al.~\citep{prestige_cultural_learning} have defined prestige bias as ``a tendency to learn from individuals to whom others have preferentially attended, learned or deferred'', and demonstrated its occurrence in in 3-4 year old children.
{Henrich and Gil-White} \citep{prestige_evolution} gave a similar definition of prestige as ``freely conferred deference'', in contrast to \emph{dominance}, and provided examples from the anthropological literature.

{Jimenez \& Mesoudi} (2019) \citep{prestigeSocialLearn} suggest that cultural evolution theory is a major factor in human ecological success. They mention that a way to acquire adaptive social information is by preferentially copying competent individuals within a valuable domain, and they call it success bias as we do in this paper. They also mention that competence within a domain is often difficult or impossible to directly asses, therefore people tend to use indirect cues of success, as suggested by {Henrich and Gil-White}.

In our paper, we examine the effect of both success and prestige biases, and their relationship in contribution to the cultural evolution of populations. We call the joint contribution of both biases \emph{Influence bias}.
This is an important distinction, as prestige is a context-dependent bias, rather than a content-{dependent} bias: it does not depend on the phenotype itself but rather on the number of copiers that have already copied each role-model, which may be easier and more accurate to estimate than success. 
Prestige bias is also frequency independent (see \Cref{cor:GBD} below), and thus it differs from conformity, which depend on the frequency of a trait in the population or in a sample of role-models, rather than the social dynamics of copying.

In contemporary human society, social media make it especially easy to estimate the social and cultural influence individuals have over others, which can have an effect on decision making. Online social networks such as \emph{Facebook} and \emph{Instagram} are known to affect the influence of individuals \citep{social_influence,social_media,influence_analysis}, and specific marketing practices have been invented to capitalize on this effect \citep{facebook_marketing}.

In the following, we develop a stochastic model of cultural transmission with influence bias that combines both cross-domain indirect success bias and prestige bias.
We develop analytic approximations for this model and analyze its dynamics. We also find approximations for the probability and time to fixation of a `successful' phenotype (i.e., that is {subject to} success bias). Comparing these approximations to Kimura's approximations for the fixation of a favorable allele \citep{kimura,kimura_average}, we demonstrate that success and influence bias play the role of natural selection and genetic drift, respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Models}
We begin with a continuous trait model with indirect bias, previously suggested by Boyd and Richerson~\citep{evolutionBook}. We extend this model to include influence bias. We then develop a dichotomous trait model. Note that the indirect bias is due to an indirect evaluation, in which a certain phenotype is used to evaluate the success of potential role-models.

@@

\paragraph{Prestige bias.}
Here we introduce a new element to the model by assuming that in each generation, copiers choose their role-models one by one.
We formulate this assumption in the following.
{Denote by} $K_{i,j}$  the number of copiers that {choose} role-model $j$ after copier $i$ chose a role-model. Thus, $i$ out of $N$ copiers {had} already {chosen} a role-model, $\sum_{j=1}^N{K_{i,j}} = i$, and there are $N-i$ copiers that have yet to choose a role-model.
The stochastic process of role-model choice, 
\begin{equation} \label{eq:process}
\big\{\vec{K}_i = (K_{i,1}, \ldots, K_{i,N}) \big\}_{i=1}^N \;,
\end{equation}
is described by the recurrence equation
\begin{equation} \label{eq:recurrence}
K_{i,j} = K_{i-1,j} + S_{i,j}, \quad i,j=1,2,\ldots,N \;,
\end{equation}
where $S_{i,j}=1$ if the $i$-th copier chose role-model~$j$ and 0 otherwise, and the initial state is $K_{0,j}=0$.

Following \cref{eq:ourF}, the probability that the $i$-th copier chose role-model $j$ is given by the influence of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:recPrestige}
Pr(S_{i,j}=1 \mid S_{1,j},S_{2,j},...,S_{i-1,j}) = G_{i,j} \;.
\end{equation}
The influence $G_{i,j}$ of role-model~$j$ in the eyes of copier~$i$ is determined by success--the estimated biased trait value $\beta(A_{i,j})$--and prestige--the number of copiers that chose role-model $j$ before copier $i$, $K_{i-1,j}$, replacing \cref{eq:boydG} with
\begin{equation}\label{eq:prestige}
G_{i,j} = \frac{\alpha_{ij} \cdot \beta(A_{i,j}) + (1-\alpha_{ij}) \cdot K_{i-1,j}}{W_i} \;,
\end{equation}
where $W_i$ is a normalizing factor that sums the numerator over all role-models ($1\le j \le N)$ to ensure $\sum_{j=1}^{N}{G_{i,j}}=1$.

Here, the bias weight $\alpha_{i,j}$ determines the relative weighting of success and prestige bias. It is a characteristic of the interaction of role-model~$j$ with copier $i$ that determines the relative significance of direct success vs. prestige in the role-model's overall influence in the eyes of the copier.
Different individuals may evaluate the importance of success and prestige differently. Additionally, we assume each role-model displays its prestige and success individually. For example, individuals with more followers but lacking skill may emphasize the number of their followers rather than their skill.
Finally, the trait of role-model $j$ estimated by copier $i$, $A_{i,j}$, remains as in \cref{eq:relativeTrait}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Fixation probability and time}
After finding that the DM distribution is a good approximation of the (within-generation) role-model choice process, we turn our attention to the (between-generation) evolutionary dynamics.
We focus on the fixation probability and conditional fixation time (conditioned on the population reaching fixation) of a favored phenotype, using a diffusion-equation approximation approach, similar to analyses of population-genetic models \citep{kimura,kimura_average,otto_fixation}.
We are mainly interested in the effect of the bias weight, $\alpha$, which determines the relative effect of success and prestige on influence bias, given by \cref{eq:prestige}.
For simplicity, we use the dichotomous model, and we do not include role-model estimation error in this analysis, i.e $e_i=0$ for every copier $i$.
As shown above, transmission in our model is approximately DM distributed (\Cref{cor:dirichlet} and \cref{eq:copier_choose}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
%During cultural transmission, cultural traits such as attitudes, values, beliefs, and behavioral patterns are transmitted between individuals, for example via copying and social learning.
{Some} cultural traits or cultural role-models may be copied more often {than others} due to transmission biases. 
{One such} bias is success bias, in which copiers are more likely to copy a successful role-model. {Although many} models assume that success can be accurately estimated{, it} has been suggested that because it is hard to estimate success, a more common bias is \emph{prestige bias}---a bias towards role-models perceived to be successful.
This perceived success can be determined by performance with respect to another trait (indirect) \citep{evolutionBook,fijian_social_bias}, or by the influence an individual has on others \citep{prestige_cultural_learning,prestige_evolution}.

We developed a cultural-evolution model with influence bias that includes both indirect success and prestige biases, where the latter is a bias towards role-models with many copiers.
We model these biases using a stochastic role-model choice process: each copier, in {turn}, randomly chooses a role-model, and this choice is affected both by the estimated success of each potential role-model and the number of copiers that already chose each role-model (eq.~\ref{eq:prestige}).

@@

Analyzing the model with the DM distribution, we found approximations for the fixation probability and fixation time of a cultural trait under biased transmission in a constant environment.
Our approximations are similar to Kimura's evolutionary-genetic approximations, {in} that (i) the difference between the resident and invading cultural trait values, $1-\beta(A)$, is equivalent to the selection coefficient in favor of a beneficial allele, $s$, and (ii) increasing the relative weight of influence versus success bias, $\alpha$, decreases the effective population size, $N_e$~(\Cref{fig:var_alpha}).

@@

We also examined a scenario in which copiers can adapt their success-weight bias, $\alpha$, to minimize their copying error, i.e., copy trait values closer to the optimal value.
We found that as the role-model choice process proceeds (that is, more copiers make their choices), both the success-bias weight ({chosen} by copiers) and the estimation error decrease. 
The latter is significantly lower {than in} a population using a constant, fixed success-bias weight, regardless of the value of the constant weight (\Cref{fig:influence_advantage}).
This suggests that the later a copier makes its choice, the more it should rely on choices of previous copiers, and the less it should rely on its own estimation.
The rationale, then, is that the more information a copier has, e.g., by using others as information sources, the more informative and effective his choice can be.

%\paragraph{Empirical evidence of prestige bias}
Chudek et al.~\citep{prestige_cultural_learning} report the first direct tests in children that suggest the existence of prestige bias, defined as the tendency to learn from individuals to whom others have preferentially attended, learned, or deferred.
Their definition of prestige is similar to ours. They showed that the odds of 3-4 years-old children learning from an adult role-model to whom bystanders had previously preferentially attended for 10 seconds were more than twice those of learning from a role model whom bystanders ignored.
They also note that prestige effects are domain sensitive: they found that prestigious role-models were attended more when demonstrating artifact use, whereas role-models presenting food preferences had less attendants, suggesting that the domain itself (artifact use vs. food preference) can affect the attendance, and hence the prestige of the role-model.
This led to the suggestion that when the trait is costly to learn individually, prestige will have a stronger bias \citep{prestige_cultural_learning}.
It would be interesting to include costs in our model to try and observe these effects and dynamics in a large population.

@@

Prestige bias can help to cheaply estimate and acquire knowledge, which may facilitate survival and reproduction. However, it is not always the case, and there could be negative repercussions to this bias, such as invasion of maladaptive traits.
{Takahashi and Ihara} \citep{best_of_k} mention that social learning not only takes the form of random copying of other individuals, but also involves learners' choice of what to learn and from whom to learn. They suggest a best-of-k model where an individual samples~$k$ role-models and chooses the one he deems most "successful". They {mention} that a previous mathematical analysis has shown that it may sometimes result in maladaptive cultural evolution when the payoffs associated with cultural variants vary stochastically. In such a case, learners may be selectively disfavored and in the long run replaced by unbiased learners, who simply copy someone chosen at random. They developed new mathematical models that are simpler and mathematically tractable. They found that best-of-k learning, unlike unbiased learning, can facilitate the invasion of an on average inferior variant that sometimes gives a very high payoff {(see Fogarty et al.\ \citep{Fogarty2017} and references there)}. Our model, which includes influence bias, is consistent with this claim. When a maladaptive trait is ``piggybacking'' {on} a role-model with high influence, {the former} could spread in the population.
In addition, {best}-of-k learning can be stable against invasion by unbiased learning if social learning is sometimes combined with individual learning \citep{best_of_k}. 
Our model {includes only} social learning, and not individual learning, but it could be interesting to combine it with individual learning and see how it affects the dynamics.

Prestige bias can also accelerate reversal of harmful traditions such as child marriage and domestic violence. 
{Efferson et al.}\ \citep{harmful_traditions} suggest a \emph{spillover}  mechanism, in which an intervention affects a large enough group in a target population, so that others not included in the intervention also change their behavior.
They {find} that there are individuals who act as \emph{agents}, who are often {observed}, and therefore they are ideal targets for interventions. This is similar to influential role-models in our model, where a prestigious individual will be copied more often, and will therefore spread {their} trait faster and wider in the population.
They also suggest a way to use this {phenomenon} to change existing traditions in a population. It is very clear however, that just as it can be used to end harmful traditions, the same agents could start harmful traditions.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title 
\title{Prestige bias in cultural evolutionary dynamics}

% Authors
\renewcommand\Affilfont{\small}

\author[1,2]{Saar Egozi}
\author[2,3,$\dagger$]{Yoav Ram}

\affil[1]{School of Computer Science, Reichman University, Herzliya, Israel}
\affil[2]{School of Zoology, Faculty of Life Sciences, Tel Aviv University, Tel Aviv, Israel}
\affil[3]{Sagol School of Neuroscience, Tel Aviv University, Tel Aviv, Israel}
\affil[$\dagger$]{Corresponding author: yoav@yoavram.com}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\maketitle

\begin{abstract}
If the traits of more successful individuals are more likely to be adopted, the resulting cultural transmission is described as being success-biased. In contrast, if the traits of `prestigious' individuals--those individuals that have already been copied many times--are more likely to be adopted, this is described as prestige-biased cultural transmission. In this case prestige can be used as a convenient proxy for success.
Here, we study the the joint effect of success and prestige bias on cultural evolutionary dynamics using mathematical analysis and stochastic simulations.
Analytic approximations to the stochastic role-model choice process facilitate the mathematical analysis and reduce the computational complexity of simulations. Approximations are given to the fixation probability and the fixation time of an invading cultural trait in different environments. 
We show that success bias effectively plays the role of natural selection, whereas prestige bias effectively plays the role of genetic drift.	Prestige bias, which may be strong in highly social communities, also accelerates the evolutionary dynamics, as can be expected in a rich-get-richer process.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}

Cultural transmission of attitudes, preferences, beliefs, norms, and behaviors may combine vertical transmission, in which parents transmit to their offspring; oblique transmission, in which adults (teachers, leaders, and even strangers) transmit to unrelated offspring; and horizontal transmission, in which individuals from the same age cohort transmit to one another \citep{transmissionVectorsBook}.
It has been demonstrated that non-vertical cultural transmission can maintain maladaptive traits, which can be beneficial in changing environments \citep{changeEnv,Lehmann2009}.

Transmission biases may cause a cultural trait to have a higher rate of transmission than its frequency in the population.
\textit{Success bias} occurs when individuals prefer to copy from role models that demonstrate success in some activity, such as fishing, growing yams, using medicinal plants \citep{fijian_social_bias}, or hunting \citep{Mesoudi2008}, and it can increase the probability of learning a trait that is present in those successful individuals \citep{Borofsky2022}. 
Indeed, in a tournament between learning strategies  \citep{strategiesPaper}, most winning strategies included a mixture of success-biased social learning and individual learning, implying that success-biased learning is a good strategy, but that by itself it is not enough to best other strategies, even when success is measured accurately.
Jimenez and Mesoudi~\citep{prestigeSocialLearn} also note that a way to acquire adaptive social information is by preferentially copying competent individuals within a valuable domain (which they also call success bias). However, they claim that competence within a domain is often difficult or impossible to directly asses, and therefore people tend to use indirect cues of success.
Henrich and Broesch \citep{fijian_social_bias} have also suggested that direct assessment of success may be ``noisy, unreliable or unavailable'' and therefore copiers should also take into account indirect measures of perceived success (e.g., ``great fishermen may be chosen as role-models for growing yams''.)

Boyd and Richerson~\citep[Ch.~5]{evolutionBook} suggested that the assessment of success can be divided to three categories: \emph{direct bias}, \emph{indirect bias} and \emph{frequency-dependent bias}.
Direct bias occurs when one phenotype is more attractive than other phenotypes, and is evaluated by \emph{directly} testing the trait.
For example, an individual observing a ping-pong match can try the observed paddle grips to determine which grip is better.
Frequency-dependent bias occurs when the probability of copying a phenotype is higher or lower than the frequency of the phenotype among demonstrators. 
For example, suppose the common paddle grip is used by 60\% of the demonstrators; if the this grip is adopted by 80\% of copiers, then transmission is under positive frequency bias, also called \emph{conformity}; if it is adopted by 40\% of copiers, then transmission is under negative frequency bias, or \emph{anti-conformity} \citep{conformism}. 
The effects of conformity and anti-conformity on cultural evolution have been studied with both models \citep{anticonformity,Denton2021,conformity} and experiments \citep{negativeFrequency}.
Indirect bias occurs when a copier uses some observed phenotype to evaluate the attractiveness of a potential role-model. 
For example, an observer may copy the paddle grip of the ping-pong player who scored more points in the match, thus indirectly evaluating the grip by the points scored.
However, this may cause mismatches between the copied trait and the rest of the cultural or genetic repertoire of the individual \citep{Kolodny2022}.
Furthermore, Boyd and Richerson~\citep[Ch. 8]{evolutionBook} suggest that maladaptive traits may spread widely in a population if indirect bias is strong enough, e.g., by a runaway process caused by a cultural equivalent of sexual selection \citep{sexualSelectionBook}.
Indeed, helping behaviors can evolve due to horizontal transmission bias even without any benefit to the recipient~\citep{cooperation}.

Henrich and Gil-White~\citep{prestige_evolution} noted that ``the most skilled/knowledgeable role-models will, on average, end up with the biggest and most lavish clienteles, so the size and lavishness of a given model's clientele size (the \emph{prestige}) provides a convenient and reliable proxy for that person's information quality''.
Therefore, they predicted that skilled individuals have higher status, that people preferentially copy high-status individuals, and therefore that prestigious individuals may be influential even beyond their domain of expertise. 
They defined prestige as ``freely conferred deference'', in contrast to \emph{dominance}, and provided examples from the anthropological literature \citep{prestige_evolution} .
Similarly, the New Oxford American Dictionary defines prestige as the ``widespread respect and admiration felt for someone or something on the basis of a \emph{perception} of their achievements or quality''.
Chudek et al.~\citep{prestige_cultural_learning} have also defined prestige bias as ``a tendency to learn from individuals to whom others have preferentially attended, learned or deferred'' and demonstrated its occurrence in in 3-4 year old children.
Henrich and Broesch \citep{fijian_social_bias} have further suggested that prestige bias can, over generations, lead to cultural adaptations, and that although prestige can lead to maladaptive traits spreading in the population, it can also accelerate the spread of adaptive traits.

The distinction between success and prestige bias is important, as prestige is a context-dependent bias, rather than a content-dependent bias: it does not depend on the phenotype itself but rather on the number of copiers that have already copied each role-model, which may be easier and more accurate to estimate than success. 
Prestige bias is also frequency independent (see \Cref{cor:GBD} below), and thus it differs from conformity \citep{anticonformity,Denton2021,conformity}, which depends on the frequency of a trait in the population or in a sample of role-models, rather than the social dynamics of copying.

Prestige bias may be more common in humans than success bias \citep{complexityPaper}.
In contemporary human society, social media make it especially easy to estimate the social and cultural influence individuals have over others, which can have an effect on decision making. Online social networks such as \emph{Facebook} and \emph{Instagram} are known to affect the influence of individuals \citep{social_influence,social_media,influence_analysis}, and specific marketing practices have been invented to capitalize on this effect \citep{facebook_marketing}.
However, despite many mentions of prestige in the cultural evolution literature, there are few models of prestige bias.

Here, we develop a stochastic model of cultural transmission with both cross-domain indirect success bias and prestige bias to examine their relationship in contribution to the cultural evolution of populations.
We find analytic approximations for this model and analyze its dynamics. We also find approximations for the probability and time to fixation of a `successful' phenotype (i.e., that is subject to success bias). Comparing these approximations to Kimura's approximations for the fixation of a favorable allele \citep{kimura,kimura_average}, we demonstrate that success and prestige bias play the role of natural selection and genetic drift, respectively.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Models}
We begin with a continuous trait model with indirect success bias, previously suggested by Boyd and Richerson~\citep{evolutionBook}. We extend this model to include prestige bias. We then develop a dichotomous trait model. Note that the indirect success bias is due to an indirect evaluation, in which a certain phenotype is used to evaluate the success of potential role-models.

@@

\paragraph{Prestige bias.}
Here we introduce a new element to the model by assuming that in each generation, copiers choose their role-models one by one.
We formulate this assumption in the following.
Denote by $K_{i,j}$  the number of copiers that choose role-model $j$ after copier $i$ chose a role-model. Thus, $i$ out of $N$ copiers had already chosen a role-model, $\sum_{j=1}^N{K_{i,j}} = i$, and there are $N-i$ copiers that have yet to choose a role-model.
The stochastic process of role-model choice, 
\begin{equation} \label{eq:process}
\big\{\vec{K}_i = (K_{i,1}, \ldots, K_{i,N}) \big\}_{i=1}^N \;,
\end{equation}
is described by the recurrence equation
\begin{equation} \label{eq:recurrence}
K_{i,j} = K_{i-1,j} + S_{i,j}, \quad i,j=1,2,\ldots,N \;,
\end{equation}
where $S_{i,j}=1$ if the $i$-th copier chose role-model~$j$ and 0 otherwise, and the initial state is $K_{0,j}=0$.

Following \cref{eq:ourF}, the probability that the $i$-th copier chose role-model $j$ is given by the \emph{influence} of role-model $j$ in the eyes of copier $i$,
\begin{equation}\label{eq:recPrestige}
Pr(S_{i,j}=1 \mid S_{1,j},S_{2,j},...,S_{i-1,j}) = G_{i,j} \;.
\end{equation}
The influence $G_{i,j}$ of role-model~$j$ in the eyes of copier~$i$ is determined by success--the estimated biased trait value $\beta(A_{i,j})$--and prestige--the number of copiers that chose role-model $j$ before copier $i$, $K_{i-1,j}$, replacing \cref{eq:boydG} with
\begin{equation}\label{eq:prestige}
G_{i,j} = \frac{\alpha_{ij} \cdot \beta(A_{i,j}) + (1-\alpha_{ij}) \cdot K_{i-1,j}}{W_i} \;,
\end{equation}
where $W_i$ is a normalizing factor that sums the numerator over all role-models ($1\le j \le N)$ to ensure $\sum_{j=1}^{N}{G_{i,j}}=1$.

Here, the bias weight $\alpha_{i,j}$ determines the relative weighting of success and prestige bias. It is a characteristic of the interaction of role-model~$j$ with copier $i$ that determines the relative significance of direct success vs. prestige in the role-model's overall influence in the eyes of the copier.
Different individuals may evaluate the importance of success and prestige differently. Additionally, we assume each role-model displays its prestige and success individually. For example, individuals with more followers but lacking skill may emphasize the number of their followers rather than their skill (i.e., have lower $\alpha_{i,j}$ value).
Finally, the trait of role-model $j$ estimated by copier $i$, $A_{i,j}$, remains as in \cref{eq:relativeTrait}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Fixation probability and time}
After finding that the DM distribution is a good approximation of the (within-generation) role-model choice process, we turn our attention to the (between-generation) evolutionary dynamics.
We focus on the fixation probability and conditional fixation time (conditioned on the population reaching fixation) of a favored phenotype, using a diffusion-equation approximation approach, similar to analyses of population-genetic models \citep{kimura,kimura_average,otto_fixation}.
We are mainly interested in the effect of the bias weight, $\alpha$, which determines the relative effect of success and prestige bias, given by \cref{eq:prestige}.
For simplicity, we use the dichotomous model, and we do not include role-model estimation error in this analysis, i.e $e_i=0$ for every copier $i$.
As shown above, transmission in our model is approximately DM distributed (\Cref{cor:dirichlet} and \cref{eq:copier_choose}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
Some cultural traits or cultural role-models may be copied more often than others due to transmission biases. 
One such bias is success bias, in which copiers are more likely to copy a successful role-model. 
It has been suggested that because it is hard to estimate success, a more common bias is a bias towards role-models \emph{perceived} to be successful.
This perceived success can be determined by performance with respect to another trait, i.e., \emph{indirect success} \citep{evolutionBook,fijian_social_bias}, or by ``the amount of voluntary deference and attention received by models'' \citep{prestigeSocialLearn}, i.e., \emph{prestige} \citep{prestige_cultural_learning,prestige_evolution} (but see Chellappoo~\citep{rethinkPrestige} for a critical examination of the concept of prestige).

We developed a cultural-evolutionary model that includes both indirect success and prestige biases, where the latter is a bias towards role-models with many copiers.
We model these biases using a stochastic role-model choice process: each copier, in turn, randomly chooses a role-model, and this choice is affected both by the estimated success of each potential role-model and the number of copiers that already chose each role-model (eq.~\ref{eq:prestige}).

@@

Analyzing the model with the DM distribution, we found approximations for the fixation probability and fixation time of a cultural trait under biased transmission in a constant environment.
Our approximations are similar to Kimura's evolutionary-genetic approximations, in that (i) the difference between the resident and invading cultural trait values, $1-\beta(A)$, is equivalent to the selection coefficient in favor of a beneficial allele, $s$, and (ii) increasing the relative weight of influence versus success bias, $\alpha$, decreases the effective population size, $N_e$~(\Cref{fig:var_alpha}).

@@

Lastly, we examined a scenario in which copiers can adjust their success-weight bias, $\alpha$, to minimize their copying error, i.e., copy trait values closer to the optimal value.
We found that as the role-model choice process proceeds (that is, more copiers make their choices), both the success-bias weight (adjusted by copiers) and the estimation error decrease. 
The latter is significantly lower than in a population using a constant, fixed success-bias weight, regardless of the value of the constant weight (\Cref{fig:influence_advantage}).
This suggests that the later a copier makes its choice, the more it should rely on choices of previous copiers (prestige), and the less it should rely on its own estimation of the success of role-models.
The rationale, then, is that the more information a copier has, e.g., by using others as information sources, the more informative and effective his choice can be.

Others have analyzed models with interactions between different transmission biases.
Ze Hong~\citep{payoff} studied a model with both frequency-dependent bias (including conformist bias) and payoff bias. The analysis demonstrates that an integrated learning strategy may be superior to relying on either bias alone. Moreover, both biases are likely to contribute to the spread of variants with high payoff, and the magnitude of conformist bias affects the evolutionary dynamics.
% TODO add: 
% Ammar, Madeleine, Laurel Fogarty, and Anne Kandler. 2023. “Social Learning and Memory.” Proceedings of the National Academy of Sciences 120 (33): e2310033120. https://doi.org/10.1073/pnas.2310033120.
% Enquist, Magnus, and Stefano Ghirlanda. 2007. “Evolution of Social Learning Does Not Explain the Origin of Human Cumulative Culture.” Journal of Theoretical Biology 246 (1): 129–35. https://doi.org/10.1016/j.jtbi.2006.12.022.
% Denton, Kaleda K., Yoav Ram, and Marcus W. Feldman. 2022. “Conformity and Content-Biased Cultural Transmission in the Evolution of Altruism.” Theoretical Population Biology 143 (February): 52–61. https://doi.org/10.1016/j.tpb.2021.10.004.

Chudek et al.~\citep{prestige_cultural_learning} report the first direct tests in children that suggest the existence of prestige bias, defined as the tendency to learn from individuals to whom others have preferentially attended, learned, or deferred.
Their definition of prestige is similar to ours. They showed that the odds of 3-4 years-old children learning from an adult role-model to whom bystanders had previously preferentially attended for 10 seconds were more than twice those of learning from a role model whom bystanders ignored.
They also note that prestige effects are domain sensitive: they found that prestigious role-models were attended more when demonstrating artifact use, whereas role-models presenting food preferences had less attendants, suggesting that the domain itself (artifact use vs. food preference) can affect the attendance, and hence the prestige of the role-model.
This led to the suggestion that when the trait is costly to learn individually, prestige will have a stronger bias \citep{prestige_cultural_learning}.
It would be interesting to include costs in our model to try and observe these effects and dynamics in a large population.

@@

Prestige bias can help to cheaply estimate and acquire knowledge, which may facilitate survival and reproduction. However, it is not always the case, and there could be negative repercussions to this bias, such as invasion of maladaptive traits.
Takahashi and Ihara~\citep{best_of_k} mention that social learning not only takes the form of random copying of other individuals, but also involves learners' choice of what to learn and from whom to learn. They suggest a best-of-k model where an individual samples~$k$ role-models and chooses the one he deems most "successful". They mention that a previous mathematical analysis has shown that it may sometimes result in maladaptive cultural evolution when the payoffs associated with cultural variants vary stochastically. In such a case, learners may be selectively disfavored and in the long run replaced by unbiased learners, who simply copy someone chosen at random. They developed new mathematical models that are simpler and mathematically tractable. They found that best-of-k learning, unlike unbiased learning, can facilitate the invasion of an on average inferior variant that sometimes gives a very high payoff (see Fogarty et al.\ \citep{Fogarty2017} and references there). Our model, which includes both success and prestige bias, is consistent with this claim. When a maladaptive trait is `piggybacking' on a role-model with high influence (joint effect of success and prestige), the former could spread in the population.
In addition, best-of-k learning can be stable against invasion by unbiased learning if social learning is sometimes combined with individual learning \citep{best_of_k}. 
Our model includes only social learning, and not individual learning, but it could be interesting to combine it with individual learning and see how it affects the dynamics.

Prestige bias can also accelerate reversal of harmful traditions such as child marriage and domestic violence. 
Efferson et al.~\citep{harmful_traditions} suggest a \emph{spillover} mechanism, in which an intervention affects a large enough group in a target population, so that others not included in the intervention also change their behavior.
They find that there are individuals who act as \emph{agents}, who are often observed, and therefore they are ideal targets for interventions. This is similar to prestigious role-models in our model, which are copied more often, and will therefore spread their trait faster and wider in the population.
They also suggest a way to use this phenomenon to change existing traditions in a population.
It is very clear however, that just as it can be used to end harmful traditions, the same agents could start harmful traditions.

Dunbar~\citep{social_brains} hypothesized that larger, more complex brains can store and manage more information and in turn, this information can support the costs of a larger brain.
Following this, Muthukrishnan and Henrich \citep{collective_brains} suggested that prestige can directly affect human physical evolution. They present a concept called \emph{cultural brains}---brains that evolved primarily for the acquisition of adaptive knowledge.
They then develop a model that predicts a strong relationship between brain size and group size, because group size also provides access to more adaptive knowledge. 
They also presented the \emph{cumulative cultural brain} hypothesis, which proposes that human brains have evolved with an ability and tendency for selective, high-fidelity social learning. As part of this process, there are a variety of strategies and biases that have evolved to hone in on the most adaptive knowledge. These strategies and biases include direct and indirect cues of the popularity of cultural traits (e.g. success and prestige biases).
They suggest that one of the reasons for the extreme increase in brain size in humans is the ability to `cheaply' acquire adaptive knowledge via transmission biases such as prestige.

One path forward is an analysis of the dynamics of the optimal success-bias weight model, in which every copier chooses its $\alpha$. It would be interesting to see the if the mean estimation error and the varying weight, $\alpha^*$, converge to specific values, and how they are affected by the model parameters.
It may also be possible to relax the assumptions required for our approximations, such as homogeneous estimation error and success-bias weight.
Lastly, it would be interesting to analyze the continuous model and determine how much it differs from the dichotomous model.
